---
title: "TBSS Bernoulli Power Calculation"
author: "Catherine Mahoney"
format: pdf
editor: visual
---

## Power calculation for tree-based scan statistics with a Bernoulli probability model

This pipeline automates the calculation of power for tree-based scan statistics used in pharmacovigilance using the recently published, open-source implementation in the package `TBSS`. This version uses a Bernoulli probability model. To evaluate the power of the tree-based scan statistic, it implements a plasmode simulation procedure to introduce known elevation in relative risk, as well as a propensity matching procedure for the creation of matched cohorts, then calculate the proportion of trials in which the simulated node alerts. The pipeline uses the `targets` and `crew` packages to manage the parallel simulations and subsequent calculations.

## Input data

The specification of the precise study design, including inclusion/exclusion criteria, baseline period, and observation period are out of the scope of this pipeline. It is assumed that the user has created the cohorts according to their selected design and formatted the table as below, with a column for *id*, *treatment_group*, each relevant covariate, and a single *outcome* column with one row for each patient-outcome combination.

This vignette uses a simulated data set of treatment group, covariates, and outcomes from a subset of ICD-10-CM diagnosis codes corresponding to diseases of the respiratory system (J00-J99). The subset of codes is used for demonstration purposes to reduce run time.

### Sample input data

| id  | treatment_group | cov_1 | cov_2 | ... | outcome |
|-----|-----------------|-------|-------|-----|---------|
| 1   | 1               | 0     | 1     | ... | A047    |
| 1   | 1               | 0     | 1     | ... | J459    |
| 1   | 1               | 0     | 1     | ... | E119    |
| 2   | 0               | 1     | 0     | ... | M545    |
| 2   | 0               | 1     | 0     | ... | F321    |

### Tree-structured variable

Tree-based scan statistics leverage the hierarchical nature of coding systems such as ICD-10-CM to test for excess risk as multiple resolutions. TBSS requires that the tree be formatted with a *parent* and *child* column, as shown below:

| child | parent |
|-------|--------|
| J040  | J04    |
| J041  | J04    |
| J0410 | J041   |
| J0411 | J041   |

The function `tree_builder()`, contained in the repository, accepts a vector of codes and generates an appropriately formatted tree-structured variable.

## Pipeline steps

The example pipeline is detailed in the \_targets.R file, and each target and corresponding function or input contains comments to guide users in incorporating their own data and study design choices. Briefly, the steps in each simulation are as follows:

1.  `select_subsample()` Select treatment and comparator groups of the desired size; this example selects 1,000 treatment and 1,000 comparator patients

2.  `plasmode()` Introduce known elevation in relative risk of an outcome using plasmode simulation. This example uses a relative risk of 2. The user can decide whether to select a specific outcome in which to simulate excess risk or to specify a target incidence rate for the outcome to enrich and allow the function to selcet the outcome randomly. In this instance, a target incidence of .01 is specified.

3.  `ps_match()` Match treated and comparator patients based on a user-defined propensity score function as defined in `match_params.R`. In this example, the propensity score is computed as a logistic regression using sample baseline data corresponding to categories from the Elixhauser comorbidity index, but users should customise to their own study and may choose to evaluate differing degrees of control of potential confounding through choice of propensity score parameters. This example uses full matching with a caliper of 0.1, but the user may customise appropriately for their study design.

4.  `tbss_mask()` Apply TBSS to determine if the simulated elevation in risk is detected while preventing detection of unrelated statistical alerts.

This process is repeated many times to determine the power, calculated as the proportion of simulations wherein the simulated excess risk is detected. The example uses only 10 simulations for demonstration purposes; suitable numbers will vary by application, but 1,000 may be a good reference point. The package `crew` manages parallelisation. In this example, only two workers are used, but the user can customise to their own computer system.

## Modifying `match_params.R`

To specify the parameters for the stratification and estimation of expected outcome rates, the user must modify the parameters in the `match_params.R` script.

The sample arguments correspond to a binomial GLM with a logit link function, but the user can specify any valid arguments to `glm()` in the `match_args` list, along with the desired matching parameters to be passed to the `matchit()` function.

```{r, eval = F, echo = F}
# propensity score formula
ps_form <- treatment_group ~ gender + age + chf + carit + valv + pcd + pvd + hypunc +
  hypc + para + ond + cpd + diabunc + diabc + hypothy + rf + ld +
  pud + aids + lymph + metacanc + solidtum + rheumd + coag + obes +
  wloss + fed + blane + dane + alcohol + drug + psycho + depre 

# specify matching parameters, suggest these:
match_method <- "full"
match_distance <- "glm"
match_link <- "logit"
match_caliper <- 0.1

match_args <- list(
  formula = ps_form,
  method = match_method,
  distance = match_distance,
  caliper = match_caliper
)
```

## Contents of the `_targets.R` script

The script `_targets.R` directs the flow of work in the pipeline. Once the matching parameters in `match_params.R` have been specified, lines with a comment can be customised to required inputs and file locations for input and tree data.

```{r, results = "hide"}
library(targets)
library(crew)

tar_source() #configure metadata with matching parameters
tar_option_set(packages = c("readr", "dplyr", "MatchIt", "TBSS", "optmatch"),
               controller = crew_controller_local(workers = 2) #specify workers
)

list(
  tar_target(file, 
             "data/test_data.csv",    #location of your outcome and covariate data
             format = "file"
  ),
  tar_target(data, 
             get_data(file)
  ),
  tar_target(iteration_id, 
             1:10                #specify number of replicates
  ),
  tar_target(subsample, 
             select_n(data, 1000, 1000),    #specify sample sizes
             pattern = map(iteration_id)
  ),
  tar_target(sim_data, 
             plasmode(subsample,
                      target_outcome = NULL, #specify either outcome or incidence
                      target_inc = .01, 
                      rr=2),                 #specify relative risk
             pattern = map(subsample)
  ),
  tar_target(ps_data,
             ps_match(sim_data,
                      match_args),
             pattern = map(sim_data)
  ),
  tar_target(tree_file, 
             'data/test_tree.csv',                   #location of the tree file
             format = "file"
  ),
  tar_target(tree, 
             get_data(tree_file)
  ),
  tar_target(tbss, 
             tbss_mask(ps_data, 
                       model = "bernoulli",
                       tree), 
             pattern = map(ps_data)
  ),
  tar_target(final_vector, 
             unlist(tbss)
  ),
  tar_target(final_power, 
             print(mean(final_vector))
  )
  
)


```

## Running the pipeline

Run `tar_make()` to launch the pipeline. The output will enumerate each target and show which branches were already up-to-date. Finally, `tar_read(final_power)` will display the result of the complete power calculation. Here, TBSS had 20% power to detect an outcome with an incidence of .01 and a relative risk of 2 when there were 1,000 treated and 1,000 comparator patients.

```{r, warning=FALSE}
tar_make()
```

```{r, echo = F}
print(paste0("Final power: ", tar_read(final_power)))
```
